### 2015-04-01 cug_paper/01_introduction.tex
Paper:  Introduction section has reached 'rough draft' state

\todo[inline]{Discuss the general idea and the rationale behind it}

@@

Specifically, approximately one third of the applications that run on Titan do not use the GPU at all.\todo{Find a citation for this.}
There may be good reasons for this.  Utilizing the GPU's requires changes to the application and it's possible the developers haven't had the time to do that.  It's also possible that the type of computations the application performs simply don't lend themselves to GPU-based acceleration.  Regardless of why, it's clear that not all applications are benefiting from the available hardware.

\todo[inline]{Abrupt transition: we go from talking about hardware to talking about software...}

In most HPC applications, filesystem writes typically follow a very 'bursty' pattern. That is, there will be long periods with no file output while the application is in a compute phase, followed by a period of intense output when the application writes its checkpoint or output files. At ORNL, the filesystem is a shared resource and other users will also be putting a varying load on it. This means that the performance that a user sees at any given time will likely vary dramatically.

Having a sufficiently large write-back cache could improve applications' write performance in two ways.  First, if the cache is large enough to hold the entire write and the individual writes are far enough apart, then the application never has to wait on the filesystem.  It can write to the cache and then resume its computations while the cache drains to the filesystem.  Secondly, write performance should be more predictable. From the application's point of view, write speed would only limited by how fast data can be copied into the cache, and that is not effected by outside influences.

 The GPU memory and integer cores can be used to implement such a cache without taking resources away from the main application.

@@

This paper demonstrates a technique for providing such a write-back cache to applications. We use a synthetic benchmark to demonstrate performance without such a cache, with a cache that utilizes regular system ram and with a cache that utilizes ram on the GPU card. Performance benefits and any potential trade-offs and caveats are discussed.
\end{comment}

>>>>

In most HPC applications, filesystem writes typically follow a very 'bursty' pattern. That is, there will be long periods with no file output while the application is in a compute phase, followed by a period of intense output and no computation when the application writes its checkpoint or output files.  Also, at ORNL, the filesystem is a shared resource and other users will also be putting a varying load on it. This means that the performance that a user sees at any given time will likely vary dramatically.

Having a sufficiently large node-local write-back cache could improve applications' performance in two ways.  First, if the cache is large enough to hold an entire write and the individual writes are far enough apart, then the application never has to wait on the filesystem.  It can write to the cache and then resume its computations while the cache drains to the filesystem.  Secondly, write performance should be more predictable. From the application's point of view, write speed would only limited by how fast data can be copied into the cache, and since the cache is local to the node, it is not effected by outside influences.

@@

To start with, during 2014, approximately 50\% of the compute time on Titan was used on applications that did not use the GPU at all.\cite{oareport}
There may be good reasons for this.  Utilizing the GPU's requires changes to the application and it's possible the developers haven't had the time to do that.  It's also possible that the type of computations the application performs simply don't lend themselves to GPU-based acceleration.  Regardless of why, it's clear that not all applications are benefiting from the GPU.

Secondly, though AMD markets the processor as having 16 cores, it actually has only 8 floating-point units. As such, some applications that are floating-point intensive run with only 8 processes per node. This actually leaves 8 cores available so long as they primarily execute integer instructions.

 The unused GPU memory and integer cores can be used to implement a write-back cache without taking resources away from the main application.  This paper demonstrates a technique for providing such a cache to applications.   Runs were made on Titan using a synthetic benchmark to demonstrate performance both without such a cache and with a cache that utilizes ram on the GPU card. The results show definite performance benefits.

@@

This paper demonstrates a technique for providing such a write-back cache to applications. We use a synthetic benchmark to demonstrate performance without such a cache, with a cache that utilizes regular system ram and with a cache that utilizes ram on the GPU card. Performance benefits and any potential trade-offs and caveats are discussed.

\end{comment}
